### 2022-09-26 main/chapter_2ljets.tex
some more text on SM, minor tweaks

The main messages that this thesis reports are my contributions to the recently
published \atlas\ searches ``for new phenomena in events with two leptons,
jets, and missing transverse momentum''~\cite{atlas2022searches},
particularly to its electroweak part.
For brevity, I will call this publication $\twoljets$ and its electroweak
part $\twoljets$-electroweak.

@@

From \atlas' large and previous Run~2 dataset, our $\twoljets$ searches
have extracted a sharper image of the high energy processes that simultaneously
produce leptons, jets and missing transverse momentum in proton-proton
collisions at the Large Hadron Collider.
The core message is clear: we see no indication of any additive process
atop the Standard Model.

>>>>

\noindent The main messages that this thesis reports are my contributions to the recently
published \atlas\ searches ``for new phenomena in events with two leptons,
jets, and missing transverse momentum''~\cite{atlas2022searches},
particularly to its electroweak part.
For brevity, I will call this publication $\twoljets$ and its electroweak
part $\twoljets$-electroweak.

@@

\noindent From \atlas' large and previous Run~2 dataset, our $\twoljets$ searches
have extracted a sharper image of the high energy processes that simultaneously
produce leptons, jets and missing transverse momentum in proton-proton
collisions at the Large Hadron Collider.
The core message is clear: we see no indication of any additive process
atop the Standard Model.

### 2022-09-26 main/chapter_context.tex
some more text on SM, minor tweaks

\subsection{The Standard Model}
\label{sec:context_sm}
% \TODO{particle content figure/tables}

The Standard Model is formulated as a quantized gauge field theory with the
symmetry that its Lagrangian is invariant under local operations of the group
\begin{equation}
SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y
\end{equation}
on its fields.
Here, the subscript $C$

~\cite{
cottingham2007greenwood,
wells2020discovery,
rubakov1999classical,
thomson2013modern
}

\TODO{Higgs mechanism}

@@

\TODO{neutrinos}
\TODO{rotation between weak and mass eigenstates: PMNS matrix}

\subsubsection{Future Standard Models}

@@

Towards explaining the main result of this thesis,
a search for effects from supersymmetric models
in data from the \atlas\ detector
on the Large Hadron Collider (LHC)~\cite{
atlas2022searches,
atlas2008experiment,
lhc2008machine
},
we first some relevant facts about the design and motivations
behind \atlas.

>>>>

\subsection{The Standard Model}
\label{sec:context_sm}
% \TODO{particle content figure/tables}
The current Standard Model is characterized by three generations of fermions
(spin-$1/2$) that interact through forces mediated by vector bosons
(spin-$1$) and a scalar Higgs boson (spin-$0$).
Force particles are categorized into the strong interaction
via with massless gluons $g$,
the electromagnetic interaction via the massless photon $\gamma$,
and
the weak interaction via the very massive $W^+$, $W^-$, and $Z$ bosons.
Electroweak unification combines the weak and electromagnetic interactions
with the Higgs mechanism, which simultaneously provides the weak boson masses
and adds massive Higgs boson, whose Yukawa interactions with the fermions also
implement their masses~\cite{thomson2013modern}

Mathematically, the Standard Model is described by a quantized gauge field
theory with the symmetry group
$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
meaning that its Lagrangian is invariant under local operations of group
elements on its fields in certain representations.

Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
indicates that they operate on colour charge triplets.

Unified electroweak interactions come from $SU\!(2)_L$ and $U\!(1)_Y$, whose
four bosons $W^1$, $W^2$, $W^3$, and $B$ respectively would be massless
without the Higgs mechanism, which inserts four scalar fields and
`spontaneously breaks' the gauge symmetry with a potential that moves the
ground state of the model to a gauge-asymmetric configuration.
From this new ground state, linear combinations of the electroweak fields
acquire masses while ``eating up'' three of the four scalar fields to yield
the three weak bosons and the photon:
$(W^1, W^2)$ mix into $(W^+, W^-)$, and $(W^3, B)$ mix into $(Z, \gamma)$.
The fourth and final scalar field is the Higgs boson $h$~\cite{
rubakov1999classical,
cottingham2007greenwood
}.

The fermions of the Standard Model are listed in Table~\TODO{}.
Older Standard Models included only two fermion generations and no
neutrino masses~\cite{
wells2020discovery
},\TODO{cite more}
but a modern model must include both three generations and non-zero neutrino
masses to be compatible with experiment:
three fermion generations are directly established
\TODO{kaon mixing CPV, direct observation of tau/top/bottom, Z width},
as are neutrino masses from their flavour oscillations demonstrating that
they experience proper time~\TODO{cite}.

@@

\TODO{neutrinos}
\TODO{rotation between weak and mass eigenstates: PMNS matrix!!}

\subsubsection{Opportunities for future Standard Models}

@@

\noindent Towards explaining the main result of this thesis,
a search for effects from supersymmetric models
in data from the \atlas\ detector
on the Large Hadron Collider (LHC)~\cite{
atlas2022searches,
atlas2008experiment,
lhc2008machine
},
we first some relevant facts about the design and motivations
behind \atlas.

### 2022-09-26 main/chapter_searches.tex
some more text on SM, minor tweaks

Our task in scientific data analysis is to distil useful messages from the
data, and to report that message to the wider community.
Various messages could be extracted from given data,
and those messages will have differing values to their recipients.
Data analysis is not an automatic process, but a design with many judgements
that are informed by theory, intuition, and convention to give value to their
results.

@@

Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of randomly occurring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$ when
$N$ have satisfied $b$.
Interpreted with the Binomial model of
Equation~\ref{eqn:searches_binomial}, this definition converges to our more
general probability whenever it applies.

@@

Data analysis in modern High Energy Physics does not exactly follow frequentist
ideology, but is a hands-on, practical, and evolving ecosystem that has adapted
to avoid some of the more glaring flaws when attacking real problems.
We have seen in Section~\ref{sec:searches_data_analysis} how likelihood ratios
can clearly express our scientific interest in how well competing models
predict data, and how those likelihoods can be manipulated in probability
algebra.
We have also seen in Section~\ref{sec:searches_frequentist} how 20th century
writers championed likelihoods not on the data themselves but on inequalities,
forming \pvalues\ and confidence intervals in reductions that throw away
information and obfuscate the results.
This section reviews how these objects are applied generally in conventional
analysis at the LHC, and specifically towards the \atlas\ search that this
thesis presents.

>>>>

\noindent Our task in scientific data analysis is to distil useful messages from the
data, and to report that message to the wider community.
Various messages could be extracted from given data,
and those messages will have differing values to their recipients.
Data analysis is not an automatic process, but a design with many judgements
that are informed by theory, intuition, and convention to give value to their
results.

@@

\noindent Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of randomly occurring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$ when
$N$ have satisfied $b$.
Interpreted with the Binomial model of
Equation~\ref{eqn:searches_binomial}, this definition converges to our more
general probability whenever it applies.

@@

\noindent Data analysis in modern High Energy Physics does not exactly follow frequentist
ideology, but is a hands-on, practical, and evolving ecosystem that has adapted
to avoid some of the more glaring flaws when attacking real problems.
We have seen in Section~\ref{sec:searches_data_analysis} how likelihood ratios
can clearly express our scientific interest in how well competing models
predict data, and how those likelihoods can be manipulated in probability
algebra.
We have also seen in Section~\ref{sec:searches_frequentist} how 20th century
writers championed likelihoods not on the data themselves but on inequalities,
forming \pvalues\ and confidence intervals in reductions that throw away
information and obfuscate the results.
This section reviews how these objects are applied generally in conventional
analysis at the LHC, and specifically towards the \atlas\ search that this
thesis presents.
